---
title: "Genoscape-cawa"
author: "Caitlin Miller"
date: "11/23/2022"
output: html_document
---


```{r setup, include=T}
#BiocManager::install("SNPRelate")
#remotes::install_github("eriqande/TESS3_encho_sen")  # for special version of tess3r
#remotes::install_github("eriqande/genoscapeRtools")  # for Eric's genoscapeRtools
#install.packages(c("raster", "sf", "tidyverse", "fields", "downloader"))
library(raster)  # important to load before tidyverse, otherwise it masks select()
library(tidyverse)
library(readr)
library(sf)
library(ggspatial)
library(ggplot2)
library(readxl)
library(raster)
library(rgdal)
library(smoothr)
library(lwgeom)
library(rgeos)
library(readxl)
```


##Make the raster of interpolated Q values

This is using the adaptive admixture values.

```{r reading in and manipulating data}

##read in Q values from admixture
Q_tibble <- readr::read_table("../admix/adaptInput/cawa.adapt.all.221123.run1.3.Q", col_names = FALSE)
colnames(Q_tibble) <- c("Southern", "Eastern", "Northwestern")

##read in sample meta data and combine with Q data
meta <- readxl::read_excel("../cawa_seq_metadata.xlsx")
Q_tibble$sample <- meta$sample
info <- left_join(Q_tibble,meta, by="sample")

##make Q value matrix
q_matrix <- Q_tibble %>%
  select(-sample) %>%
  as.matrix()

##make lat/long matrix
long_lat_tibble <- info %>%
  select(long, lat)

long_lat_matrix <- long_lat_tibble %>%
  as.matrix()

##read in the shapefile for the breading range
breeding_range <- st_read("../../cawa-range-map/map/CAWA.breed_season.sf.WGS84.Ebird.shp")
##check polygons
ggplot(breeding_range) +
  geom_sf() +
  theme_bw()

rm(list=ls()[!ls() %in%c("breeding_range", "long_lat_matrix", "q_matrix", "info")])
```

#Interpolate the Q values across the range

After reading in lat/long and Q values for sequencing samples, I will interpolate the rest of the range by using krieging.

```{r interpolating and krieging across the breeding range}

#assign colors to areas
cluster_colors <-  c(
  Southern = "#ff4a6a", ##pink
  Northwestern = "#229047", ##green
  Eastern = "#2637D4" ##blue-purple
)

genoscape_brick <- tess3r::tess3Q_map_rasters(
  x = q_matrix,
  coord = long_lat_matrix,
  map.polygon = breeding_range,
  window = extent(breeding_range)[1:4],
  resolution = c(300,300), # if you want more cells in your raster, set higher
  # this next lines need to to be here, but don't do much...
  col.palette = tess3r::CreatePalette(cluster_colors, length(cluster_colors)),
  method = "map.max",
  interpol = tess3r::FieldsKrigModel(10),
  main = "Ancestry coefficients",
  xlab = "Longitude",
  ylab = "Latitude",
  cex = .4
)
# after that, we need to add names of the clusters back onto this raster brick
names(genoscape_brick) <- c("Southern" , "Eastern","Northwestern")

##edit the raster into a nicer format, where there are probabilities of population across space
##alpha values now correspond to probabilities- opaque is high prob
genoscape_rgba <- genoscapeRtools::qprob_rando_raster(
  TRB = genoscape_brick,
  cols = cluster_colors,
  alpha_scale = 2.0,
  abs_thresh = 0.0,
  alpha_exp = 1.55,
  alpha_chop_max = 230
)

# at this point, we must be explicit about adding a projection to the raster.
# This adds the info for a regular lat-long projection
crs(genoscape_rgba) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"


```

```{r writing raster brick and saving separate shapefiles}
writeRaster(genoscape_brick,file="brick/cawa.genoscape",bylayer=T,format="GTiff",suffix="numbers", overwrite=T)

files<-list.files(".", pattern='cawa.genoscape_', recursive=T)

nodelist=list()
for (file in files) {
c1=raster(file)
c1c=cut(c1,breaks=c(-Inf,0.5,Inf))
c1p=rasterToPolygons(c1c,function(j){j==2},dissolve=TRUE)
proj4string(c1p)=CRS("+proj=longlat +datum=WGS84")
c1p=st_as_sf(c1p)
c1pd=drop_crumbs(c1p,threshold=units::set_units(1000, km^2))
c1pe=fill_holes(c1pd,threshold=units::set_units(1000, km^2))
c1smooth=smooth(c1pe,method="ksmooth",smoothness=3)
c1smooth=as(c1smooth,'Spatial')
nodelist[[file]]=c1smooth
}

#allow overlap:
node1<-nodelist[[1]]
node2<-nodelist[[2]]
node3<-nodelist[[3]]


##Remove overlaps
node1a=nodelist[[1]]-nodelist[[2]]
node1b=node1a-nodelist[[3]]

node2a=nodelist[[2]]-nodelist[[1]]
node2b=node2a-nodelist[[3]]

node3a=nodelist[[3]]-nodelist[[1]]
node3b=node3a-nodelist[[2]]

#Write out shapefiles for each. Before this, you might need to add the projection (somehow is lost)-- you can see below I read in the shapefile and added it, but if you can do before writing out below, you'd be better off! 

#If have 3 nodes and allow overlap
writeOGR(node1, dsn="brick/",layer = paste0(substr(files[[1]],1,22),".5ovlp",sep=""), driver = "ESRI Shapefile")
writeOGR(node2, dsn="brick/",layer = paste0(substr(files[[2]],1,22),".5ovlp",sep=""), driver = "ESRI Shapefile")
writeOGR(node3, dsn="brick/",layer = paste0(substr(files[[3]],1,22),".5ovlp",sep=""), driver = "ESRI Shapefile")

##If have 3 nodes and do not allow overlap
writeOGR(node1b, dsn="brick/",layer = paste0(substr(files[[1]],1,22),".5no_ovlp",sep=""), driver = "ESRI Shapefile")
writeOGR(node2b, dsn="brick/",layer = paste0(substr(files[[2]],1,22),".5no_ovlp",sep=""), driver = "ESRI Shapefile")
writeOGR(node3b, dsn="brick/",layer = paste0(substr(files[[3]],1,22),".5no_ovlp",sep=""), driver = "ESRI Shapefile")

brick1 <- st_read("brick/brick_cawa.genoscape_1.5no_ovlp.shp") %>% st_set_crs(4326) %>% st_transform(crs=4326)
brick2 <- st_read("brick/brick_cawa.genoscape_2.5no_ovlp.shp") %>% st_set_crs(4326) %>% st_transform(crs=4326)
brick3 <- st_read("brick/brick_cawa.genoscape_3.5no_ovlp.shp") %>% st_set_crs(4326) %>% st_transform(crs=4326)

#old bricks
brick1 <- st_read("../gradFor/au-shapes/CAWA.genoscape_brick_South.5no_ovlp2.shp") %>% st_set_crs(4326) %>% st_transform(crs=4326)
brick2 <- st_read("../gradFor/au-shapes/CAWA.genoscape_brick_Cen.5no_ovlp2.shp") %>% st_set_crs(4326) %>% st_transform(crs=4326)
brick3 <- st_read("../gradFor/au-shapes/CAWA.genoscape_brick_NW.5no_ovlp2.shp") %>% st_set_crs(4326) %>% st_transform(crs=4326)
pdf("brick/old.bricks.cawa.breeding_polygons.5.no_overlap.pdf")
ggplot() +
  geom_sf(data = brick1, fill = "green4", color = "green4") +
  geom_sf(data = brick2, fill = "blue", color = "blue") +
  geom_sf(data = brick3, fill = "red", color = "red") +
  theme_bw()
dev.off()


```
#Getting prettier maps

Now that the genoscape brick is formed, I can make some pretty maps.

```{r making genoscape maps}
##Adding sampling locations
ggplot() +
  layer_spatial(genoscape_rgba) +
  geom_spatial_point(
    data = info, 
    mapping = aes(x = long, y = lat), size = 1
  ) +
  theme_bw() +
  coord_sf()


##extent of breeding range
domain <- c(
  xmin = -140,
  xmax = -50,
  ymin = 0,
  ymax = 75
)

##Get coastlines
coast_cropped <- st_read("../../cawa-range-map/ne_shapefiles/ne_10m_coastline.shp") %>% st_crop(domain)
##Get countries
countries_cropped <-  st_read("../../cawa-range-map/ne_shapefiles/ne_10m_admin_0_boundary_lines_land.shp") %>% st_crop(domain)
##Get states and provinces
states_cropped <- st_read("../../cawa-range-map/ne_shapefiles/ne_10m_admin_1_states_provinces_lines.shp") %>% st_crop(domain)
##Get lakes
lakes <- st_read("../../cawa-range-map/ne_shapefiles/ne_10m_lakes.shp")



mapg <- ggplot() +
  geom_sf(data = coast_cropped, fill = NA) +
  geom_sf(data = countries_cropped, fill = NA) +
  geom_sf(data = lakes, fill = NA) +
  ggspatial::layer_spatial(genoscape_rgba) +
  theme_bw()

# now plot it under default lat-long projection
mapg + coord_sf()

```


#Making a terrain background projection and rectangular map

Adding terrain because it is nice to have and using a slightly less distorted looking coordinate system

```{r terrrain maps}
##add the terrain background and project

##projection change
lamproj <- "+proj=lcc +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-100 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"

##extent of breeding range, plus a little extra for cushion
domain <- c(
  xmin = -140,
  xmax = -50,
  ymin = 25,
  ymax = 65
)
##load in pretty terrain map
hypso_cropped <- brick("geospatial/HYP_HR_SR_OB_DR/HYP_HR_SR_OB_DR.tif") %>% crop(extent(domain))

big_one <- ggplot() +
  ggspatial::layer_spatial(hypso_cropped) +
  geom_sf(data = countries_cropped, fill = NA, size = 0.15) +
  geom_sf(data = states_cropped, fill = NA, size = 0.1) +
  ggspatial::layer_spatial(genoscape_rgba) +
  theme_bw() +
  coord_sf(crs = st_crs(lamproj))

svg(file="output/large_geno.svg")
big_one
dev.off()
# Here are lat longs for trectangle limits
pts <- tribble(
  ~place, ~long, ~lat,
  "Atlanta", -84.4230019, 33.7489905,
  "WQC", -59.2808027, 47.59458,
  "Tungsten",  -128.3029719, 61.9496199
)

# here we turn that into an sf object
pts_sf <- st_as_sf(pts, coords = c("long", "lat"),
                   crs = 4326)  # 4326 is shorthand for "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

# and now we transform it to lambert
pts_lambert <- st_transform(pts_sf, crs = lamproj)
##print and choose outermost x/y
st_as_text(pts_lambert$geometry)

rectangled <- ggplot() +
  ggspatial::layer_spatial(hypso_cropped) +
  geom_sf(data = countries_cropped, fill = NA, size = 0.15) +
  geom_sf(data = states_cropped, fill = NA, size = 0.1) +
  ggspatial::layer_spatial(genoscape_rgba) +
  theme_bw() +
  coord_sf(
    crs = st_crs(lamproj),
    xlim = c(-1481929, 2791740),
    ylim = c(-531897.9, 2591308),
    expand = FALSE)  # expand = FALSE means put the axes right at the xlim and ylim
svg("output/rectangle_geno.svg")
rectangled
dev.off()
```

# Cropping and simplifying map background

```{r interpolating and krieging across the breeding range for adaptive}
rm(list=ls()[!ls() %in%c("genoscape_rgba", "breeding_range")])

us <- ggplot2::map_data(map = "usa")
can <- ggplot2::map_data(map = "world", region = "canada")
big_one <- ggplot() +
  #ggspatial::layer_spatial(hypso_cropped) +
  geom_polygon(data = us, aes(x = long, y = lat, group = group), fill = NA, color = "gray") +
    geom_polygon(data = can, aes(x = long, y = lat, group = group), fill = NA, color = "gray") +
  ggspatial::layer_spatial(genoscape_rgba) +
  geom_sf(data=breeding_range, fill=NA, color = "black") +
  theme_bw() +
  coord_sf(xlim = c(-128, -55), ylim = c(30, 61), expand = FALSE, crs=4326)

svg(file="output/large_adapt_geno_bw.svg")
big_one
dev.off()
pdf(file="output/large_adapt_geno_bw.pdf")
big_one
dev.off()

```

