---
title: "SRS_PCA"
author: "Caitlin Miller"
date: "07/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#remotes::install_github("eriqande/srsStuff", upgrade = FALSE)
library(srsStuff)
library(tidyverse)
library(readxl)
library(RColorBrewer)
```


## Covariance analysis

```{r manipulate data}
##Did this bit on cluster with runR.sbatch and srs.input.R script
# snames <- read_lines("samples1X.list")
#CAWA.filt<-srs_covar(file="copy/cawa.bqsr.miss0.2.qual30.1X.fst85.filtRda.recode.AD.txt", sample_names=snames, freq_thresh = 0)
#saveRDS(CAWA.filt, "copy/cawa.srs.filt.85fst.RDA.rds")

allCAWA <- readRDS("cawa.srs.filt.85fst.RDA.rds")
```

Incorporate the meta data

```{r}
meta<-read_excel("../cawa_seq_metadata.xlsx")
meta <- meta %>% dplyr::select(sample,group,site,sex, type,cov, plate, merged, year, platform) %>% rename(vcf_name=sample,Region_site=group,Region_pop=site, sex=sex, cov=cov,plate=plate,merged=merged, year=year, platform=platform)
meta$cov <- round(meta$cov, 0)

```

Next to create the matrix \# A function from Eric to get everything prepped and tidy

```{r prepoutput function}
prep_output <- function(A) {

  # first for each individual i, get the number of sites at which
  # that individual had a read and also another individual, j, averaged
  # over all j.  This tells us on average how many markers went into
  # compute the covariance for each individual.
  m <- A$M
  diag(m) <- 0
  ave_sites <- tibble(
    vcf_name = A$sample_names,
    ave_sites = rowMeans(m) * ncol(m) / (ncol(m) - 1)
  )

  # do the whole eigendecomposition on the standard covariance matrix.
  eig <- eigen(A$Cov)
  colnames(eig$vectors) <- sprintf("PC-%02d", 1:ncol(eig$vectors))


  pca_tib <- as_tibble(eig$vectors[,1:6]) %>%
    mutate(vcf_name = A$sample_names) %>%
    dplyr::select(vcf_name, everything())

  pca_long <- pca_tib %>%
    tidyr::gather(., key = "PC", "val", -vcf_name)

  # then expand a grid of the possible comparisons (ordered)
  pca_pairs <- expand.grid(vcf_name = pca_tib$vcf_name,
                           PCx = sprintf("PC-%02d", 1:6),
                           PCy = sprintf("PC-%02d", 1:6),
                           stringsAsFactors = FALSE) %>%
    tibble::as_tibble() %>%
    dplyr::left_join(., pca_long, by = c("vcf_name", "PCx" = "PC")) %>%
    dplyr::rename(val_x = val) %>%
    dplyr::left_join(pca_long, by = c("vcf_name", "PCy" = "PC")) %>%
    dplyr::rename(val_y = val)  %>%
    left_join(ave_sites, by = "vcf_name") %>%
    left_join(meta, by = "vcf_name")

  WaG <- (A$IBS - A$Mt_S) / (1 - A$Mt_S)
  eig <- eigen(WaG)
  colnames(eig$vectors) <- sprintf("PC-%02d", 1:ncol(eig$vectors))


  pca_tib <- as_tibble(eig$vectors[,1:6]) %>%
    mutate(vcf_name = A$sample_names) %>%
    dplyr::select(vcf_name, everything())

  pca_long <- pca_tib %>%
    tidyr::gather(., key = "PC", "val", -vcf_name)

  pcp <- expand.grid(vcf_name = pca_tib$vcf_name,
                     PCx = sprintf("PC-%02d", 1:6),
                     PCy = sprintf("PC-%02d", 1:6),
                     stringsAsFactors = FALSE) %>%
    tibble::as_tibble() %>%
    dplyr::left_join(., pca_long, by = c("vcf_name", "PCx" = "PC")) %>%
    dplyr::rename(WaG_x = val) %>%
    dplyr::left_join(pca_long, by = c("vcf_name", "PCy" = "PC")) %>%
    dplyr::rename(WaG_y = val)

  # then join those both and return them
  left_join(pcp, pca_pairs,
            by = c("vcf_name", "PCx", "PCy"))
}
```

##CAWA Breeding Site Analysis

```{r get data arranged and colors sorted}
breedingCAWA.prep <- prep_output(allCAWA)
factor.levels <- sort(c(unique(meta$Region_pop)))
                   
breedingCAWA.prep$Region_pop <- factor(x = breedingCAWA.prep$Region_pop,
                                   levels = factor.levels)

breedingCAWA.prep <- arrange(breedingCAWA.prep, Region_pop)

##Save colors to sites that are close to each other
##Western blues 
AB.cols <- brewer.pal(n = 9, name = "Blues")[c(4:6)]
##Eastern oranges
MB.col <-brewer.pal(n = 9, name = "Oranges")[2]
##Eastern oranges
ME.col <-brewer.pal(n = 9, name = "Oranges")[3]
##Eastern oranges
MN.col <-brewer.pal(n = 12, name = "Set3")[4]
##Eastern oranges
NB.col<-brewer.pal(n = 11, name = "PuOr")[1]
##Pink
NC.col<-brewer.pal(n = 8, name = "Dark2")[4]
##Eastern oranges
NH.col <-brewer.pal(n = 9, name = "Oranges")[4]
##Eastern oranges
NY.col <-brewer.pal(n = 9, name = "Oranges")[5]
##Eastern oranges
ON.col <-brewer.pal(n = 9, name = "Oranges")[5]
##Southern greens
PA.cols <- brewer.pal(n = 9, name = "BuGn")[c(6:7)]
##Eastern oranges
QC.cols <-brewer.pal(n = 9, name = "Oranges")[c(6,7)]
##Eastern oranges
RI.col<-brewer.pal(n = 9, name = "Oranges")[8]
##Eastern oranges
WI.col<-brewer.pal(n = 11, name = "RdGy")[3]
##Southern greens
WV.col<-brewer.pal(n = 9, name = "BuGn")[4]


##Save colors
pca.color <- c(AB.cols, MB.col, ME.col, MN.col, NB.col, NC.col, NH.col, NY.col, ON.col, PA.cols, QC.cols, RI.col, WI.col, WV.col)

```

##Initial Plots Facet grid:

```{r graph PCs by different potential influences}
subsetBreedingCAWA.prep <- breedingCAWA.prep %>% filter(PCx == c("PC-01", "PC-01", "PC-03")) %>%  filter(PCy == c( "PC-02", "PC-03"))


bp <- ggplot(subsetBreedingCAWA.prep, aes(x = val_x, y = val_y, color = Region_pop))+
  facet_grid(PCy ~ PCx) +
  scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(legend.key.size = unit(0.1, "cm"))
  
bp

isotopeGroup <- ggplot(subsetBreedingCAWA.prep, aes(x = val_x, y = val_y, color = as.factor(Region_site)))+
  facet_grid(PCy ~ PCx) +
  scale_color_manual(values = c("green", "orange", "purple")) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(legend.key.size = unit(0.1, "cm"))
  
isotopeGroup

sex <- ggplot(subsetBreedingCAWA.prep, aes(x = val_x, y = val_y, color = sex))+
  facet_grid(PCy ~ PCx) +
  #scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(legend.key.size = unit(0.1, "cm"))
  
sex

plate <- ggplot(subsetBreedingCAWA.prep, aes(x = val_x, y = val_y, color = as.factor(plate)))+
  facet_grid(PCy ~ PCx) +
  #scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(legend.key.size = unit(0.1, "cm"))
  
plate

platform <- ggplot(subsetBreedingCAWA.prep, aes(x = val_x, y = val_y, color = platform))+
  facet_grid(PCy ~ PCx) +
  #scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(legend.key.size = unit(0.1, "cm"))
  
platform

type <- ggplot(subsetBreedingCAWA.prep, aes(x = val_x, y = val_y, color = type))+
  facet_grid(PCy ~ PCx) +
  #scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(legend.key.size = unit(0.1, "cm"))
  
type

year <- ggplot(subsetBreedingCAWA.prep, aes(x = val_x, y = val_y, color = year))+
  facet_grid(PCy ~ PCx) +
  #scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(legend.key.size = unit(0.1, "cm"))
  
year

cov <- ggplot(subsetBreedingCAWA.prep, aes(x = val_x, y = val_y, color = as.factor(cov)))+
  facet_grid(PCy ~ PCx) +
  #scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(legend.key.size = unit(0.1, "cm"))
  
cov
```

##Figure out which indivduals, if any, to remove

```{r find the covariance of all inds}
cov_and_M <- function(A) {
  tibble(vcf_name_1 = rep(A$sample_names, times = nrow(A$Cov)),
         vcf_name_2 = rep(A$sample_names, each = ncol(A$Cov)),
         covar = as.vector(A$Cov),
         M = as.vector(A$M),
         WaG = as.vector((A$IBS - A$Mt_S) / (1 - A$Mt_S))
  )
}

CM <- cov_and_M(allCAWA)

CM_cut <- CM %>%
  filter(vcf_name_1 < vcf_name_2) %>%
  arrange(desc(M)) %>%
  mutate(m_group = cut(M, breaks = c(0, 1e6, 5e6, 1e7, 1e8, 2e8)))


g <- ggplot(CM_cut, aes(x = covar, y = 2, colour = log10(M), size = 3)) +
  geom_jitter(width = 0, height = 1) +
  scale_colour_viridis_c() +
  facet_wrap(~m_group, ncol = 1) +
  theme_bw()

g
```

Save file.

```{r save the plot of covariance}
ggsave(g, filename = "plots/cawa-filt85fstRDA-missing-covar-and-M-facets.pdf",
      width = 20, height = 20)
```

##Create table to individuals to remove. Choose the covariance cut off that works for your data. I choose \>0.1

```{r Find individuals with covariance >0.1}
outlier_CM_cut <- CM_cut %>% 
  filter(covar > 0.1)
outlier_covar <- unique(c(unique(outlier_CM_cut$vcf_name_1), unique(outlier_CM_cut$vcf_name_2)))


meta[meta$vcf_name %in% outlier_covar,]
```

There are no individuals to remove with covariance \>0.1, so I will not change anything

## PCs 1 and 2

```{r check PC1 and PC2}
pc.eigenvalues <- eigen(allCAWA$Cov)$values
pc.var.expl <- pc.eigenvalues[1:2] / sum(pc.eigenvalues) * 100

PC12.prep <- breedingCAWA.prep %>%
  filter(PCx == "PC-01", PCy == "PC-02")

pca12 <- ggplot(PC12.prep, aes(x = val_x, y = val_y, color = Region_pop)) +
  scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(aspect.ratio=1) + 
  theme(legend.key.size = unit(0.1, "cm"))+
  xlab("PC1: 0.995%") +
  ylab("PC2: 0.934%")
  
pca12
```

Saving the plots for later

```{r save the file of the PCA}
ggsave(pca12, filename = "plots/cawa.filt.85fst.RDA.PC1.PC2.pdf",
      width = 7, height = 7)

svg("plots/cawa.filt.85fst.RDA.PC1.PC2.svg")
pca12
dev.off()
```

##PC2 and PC3

```{r check PC2 and PC3}
pc.eigenvalues <- eigen(allCAWA$Cov)$values
pc.var.expl <- pc.eigenvalues[2:3] / sum(pc.eigenvalues) * 100

PC23.prep <- breedingCAWA.prep %>%
  filter(PCx == "PC-02", PCy == "PC-03")

pca23 <- ggplot(PC23.prep, aes(x = val_x, y = val_y, color = Region_pop)) +
  scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(aspect.ratio=1) + 
  theme(legend.key.size = unit(0.1, "cm"))+
  xlab("PC2: 0.934%") +
  ylab("PC3: 0.891%")
  
pca23
```

Save the plots

```{r save the file of the PCA2-3}
ggsave(pca23, filename = "plots/cawa.filt.85fst.RDA.PC2.PC3.pdf",
      width = 7, height = 7)

svg("plots/cawa.filt.85fst.RDA.PC2.PC3.svg")
pca23
dev.off()
```

## PCs 1 and 2

```{r save PC1 and PC2 with shape coding instead of color}
pc.eigenvalues <- eigen(allCAWA$Cov)$values
pc.var.expl <- pc.eigenvalues[1:2] / sum(pc.eigenvalues) * 100

PC12.prep <- breedingCAWA.prep %>%
  filter(PCx == "PC-01", PCy == "PC-02")
PC12.prep <- arrange(PC12.prep, Region_pop)

##Save colors
pca.color <- c("gray", "gray", "purple")
pca.shapes <- c("0", "1", "2")

pca12shape <- ggplot(PC12.prep, aes(x = val_x, y = val_y, color = Region_site, shape =Region_site)) +
  scale_color_manual(values = pca.color) +
 geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(aspect.ratio=1) + 
  theme(legend.key.size = unit(0.1, "cm"))+
  xlab("PC1: 0.995%") +
  ylab("PC2: 0.934%")
  
pca12shape
```

Save the plot

```{r save the file of the PCA1-2 with shapes}
ggsave(pca12shape, filename = "plots/cawa.filt.85fst.RDA.PC1.PC2.shapes.pdf",
      width = 7, height = 7)

svg("plots/cawa.filt.85fst.RDA.PC1.PC2.shapes.svg")
pca12shape
dev.off()
```